Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2021-04-05 mice_synthesizing/paper.Rmd
Update imputation how to

The remainder of this paper is organized as follows. First, we will shortly describe the `mice` algorithm for the creation of synthetic data. The aim is to generate synthetic sets that reassure the privacy and confidentiality of the participants. Second, we will demonstrate a straightforward workflow for imputation of synthetic data with `mice`. The third step is to summarize the variance estimators that have been proposed in the past. Finally, we will assess which variance estimator provides valid inferences from the synthetic data.

@@

The `mice` package has been developed for multiple imputation of missing data. In this context, the aim is to replace missing values due to nonresponse by plausible values from the posterior predictive distribution of the variable containing the missings. Doing so, `mice` makes use of fully conditional specification [FCS; @vanbuuren_fully_2006], which breaks down the multivariate distribution of the data $\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})$ into $j = 1, 2, \dots, k$ univariate conditional densities, with $k$ the number of variables in the data. Using FCS, for every variable containing missing values a model is constructed, to impute the missing values $Y_{j, mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ on a variable by variable basis. Note that the predictor matrix $Y_{-j}$ may contain yet imputed values from an earlier imputation step, and thus will be updated after every iteration. This procedure is applied $m$ times, resulting in $m$ completed datasets $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})$. __MAYBE ADD AN EXAMPLE OF HOW TO DO IT IN `mice`__

This approach can be straightforwardly extended to a synthetic data framework. However, rather than imputing missing data, we overimpute actually observed data, that is, we replace observed values by multiple imputations. For simplicity, assume that the data is completely observed (i.e., $\textbf{Y} = \textbf{Y}_{obs}$). Following the notation of @reiter_raghunathan_multiple_2007, let $Z_j = 1$ if any of the values of unit $j$ are to be replaced by imputations, and $Z_j = 0$ otherwise, with $\textbf{Z} = (Z_1, Z_2, \dots, Z_n)$. Accordingly, the data consists of values that are to be replaced and values that are to be kept (i.e., $\textbf{Y}^{(l)} = (\textbf{Y}^{(l)}_{rep}, \textbf{Y}_{nrep})$. Note the superscript $^{(l)}$ indicating the $l$th synthetic data set, with $l = 1, 2, \dots, m$. Now, instead of imputing $\textbf{Y}^{(l)}_{mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ as in the missing data case, we impute $\textbf{Y}^{(l)}_{rep}$ from the posterior distribution of $(Y_{rep} | \textbf{Y}, \textbf{Z}, \theta)$. This process results in the synthetic data $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$.

@@

Choosing an adequate method to impute the data is paramount, as a flawed imputation model will drastically reduce the data utility in further inferences. Imputation models should be as flexible as possible to capture most of the patterns in the data, and to model, possibly unanticipated, data characteristics [@murray_multiple_2018; @rubin_18years_1996]. Parametric methods, albeit easy to implement in practice, may be too restrictive to capture generally complex patterns in the data, especially in the case of nonlinear relations and interactions between multiple variables. Classification and regression trees [CART; @breiman_cart_1984] allow to model more complex patterns in the data, and have therefore been suggested as an appropriate imputation method [@reiter_cart_2005; @burgette_reiter_cart_2010; @doove_buuren_recursive_2014]. Loosely speaking, CART sequentially splits the predictor space into non-overlapping regions in such a way that the within-region variance is as small as possible after every split. As such, CART does not impose any parametric distribution on the data, making it a widely applicable method that allows for a large variety of relationships within the data [@islr_2013]. Given these appealing characteristics and the call for the use of flexible methods when multiply imputing data, we illustrate the performance of `mice` with CART as the specified imputation method.

@@

Additional parameters that come with the use of CART are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both restrain the flexibility of the imputation model. The values of these parameters ought to adhere to the call for imputation models that are as flexible as possible. Hence, the complexity parameter fixed at `cp = 1e-08` rather than the default `1e-08` and the minimum number of observations in each terminal node set at `minbucket = 3` rather than the default 5. The bias and variance of the final estimates justified these decisions. Additionally, the number of imputations was set to $m = 5$.

>>>>

The remainder of this paper is organized as follows. First, the `mice` algorithm for the creation of synthetic data will be shortly explained. The aim is to generate synthetic sets that reassure the privacy and confidentiality of the participants. Second, a straightforward workflow for imputation of synthetic data with `mice` will be demonstrated. The third step is to summarize the variance estimators that have been proposed in the past. Finally, it will be assessed which variance estimator provides valid inferences from the synthetic data.

@@

The `mice` package has been developed for multiple imputation of missing data. In this context, the aim is to replace missing values due to nonresponse by plausible values from the posterior predictive distribution of the variable containing the missings. Doing so, `mice` makes use of fully conditional specification [FCS; @vanbuuren_fully_2006], which breaks down the multivariate distribution of the data $\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})$ into $j = 1, 2, \dots, k$ univariate conditional densities, with $k$ the number of variables in the data. Using FCS, for every variable containing missing values a model is constructed, to impute the missing values $Y_{j, mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ on a variable by variable basis. Note that the predictor matrix $Y_{-j}$ may contain yet imputed values from an earlier imputation step, and thus will be updated after every iteration. This procedure is applied $m$ times, resulting in $m$ completed datasets $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$, with $\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})$.

This approach can be straightforwardly extended to a synthetic data framework. However, rather than imputing missing data, the actually observed data is overimputed, that is, the observed values are replaced by multiple imputations. For simplicity, assume that the data is completely observed (i.e., $\textbf{Y} = \textbf{Y}_{obs}$). Following the notation of @reiter_raghunathan_multiple_2007, let $Z_j = 1$ if any of the values of unit $j$ are to be replaced by imputations, and $Z_j = 0$ otherwise, with $\textbf{Z} = (Z_1, Z_2, \dots, Z_n)$. Accordingly, the data consists of values that are to be replaced and values that are to be kept (i.e., $\textbf{Y}^{(l)} = (\textbf{Y}^{(l)}_{rep}, \textbf{Y}_{nrep})$. Note the superscript $^{(l)}$ indicating the $l$th synthetic data set, with $l = 1, 2, \dots, m$. Now, instead of imputing $\textbf{Y}^{(l)}_{mis}$ with draws from the posterior predictive distribution of $P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)$ as in the missing data case, $\textbf{Y}^{(l)}_{rep}$ is imputed from the posterior distribution of $(Y_{rep} | \textbf{Y}, \textbf{Z}, \theta)$. This process results in the synthetic data $\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})$.

@@

Choosing an adequate method to impute the data is paramount, as a flawed imputation model will drastically reduce the data utility in further inferences. Imputation models should be as flexible as possible to capture most of the patterns in the data, and to model, possibly unanticipated, data characteristics [@murray_multiple_2018; @rubin_18years_1996]. Parametric methods, albeit easy to implement in practice, may be too restrictive to capture generally complex patterns in the data, especially in the case of nonlinear relations and interactions between multiple variables. Classification and regression trees [CART; @breiman_cart_1984] allow to model more complex patterns in the data, and have therefore been suggested as an appropriate imputation method [@reiter_cart_2005; @burgette_reiter_cart_2010; @doove_buuren_recursive_2014]. Loosely speaking, CART sequentially splits the predictor space into non-overlapping regions in such a way that the within-region variance is as small as possible after every split. As such, CART does not impose any parametric distribution on the data, making it a widely applicable method that allows for a large variety of relationships within the data [@islr_2013]. Given these appealing characteristics and the call for the use of flexible methods when multiply imputing data, the performance of `mice` with CART as the specified imputation method will be illustrated.

Additional parameters that come with the use of CART are the complexity parameter `cp` and the minimum number of observations in any terminal node `minbucket`, that both restrain the flexibility of the imputation model. Appropriate values for these parameters, as well as the input for the `predictorMatrix`, depend on the data at hand. Hence, the specification of these parameters will be illustrated using an actual dataset. The `boys` data [@fredriks_boys_2000] that is in-built within the `mice` package is used to illustrate the specification of these parameters accordingly. **DETERMINE WHETHER THIS DATASET SUFFICES; BECAUSE IMPUTATION MIGHT MAKE THE DATA ARTIFICIALLY SIMILAR BEFORE SYNTHESIZING, MAKING SYNTHESIZING TOO EASY** Unfortunately, this dataset does not differ from the vast majority of collected datasets, in the sense that it suffers from missing data as well. For simplicity, the data is completed using the default `mice` imputation model to create a single, fully observed dataset. The data consists of the values of $`r nrow(mice::boys)`$ Dutch boys on the following $`r ncol(mice::boys)`$ variables: age (in years), height (in cm), weight (in kg), body mass index, heid circumference (in cm), Genital Tanner stage (G1-G5), testicular volume (in ml) and region.

The values of the parameters `cp` and `minbucket` ought to adhere to the call for imputation models that are as flexible as possible. Accordingly, the complexity parameter is fixed at `cp = 1e-08` rather than the default value `1e-08` and the minimum number of observations in each terminal node is set at `minbucket` $= 3$ rather than the default value $5$. The bias and variance of the final estimates justified these decisions. The predictor matrix was specified in such a way that 
Additionally, the number of imputations was set to $m = 5$.

@@
